---
title: "Data Import"
output: pdf_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```


This document will show how to import data. 

## Import the FAS Litters CSV
```{r}
litters_df = read_csv("data/FAS_litters.csv")

litters_df = janitor::clean_names(litters_df)
```


## Look at the dataset
```{r}
litters_df

head(litters_df)
tail(litters_df, 10)
```

```{r eval = FALSE}
view(litters_df)
```


## Learning Assessment
First load the FAS_pups.csv file using the relative path

```{r}
pups_df = read_csv("data/FAS_pups.csv")

pups_df = janitor::clean_names(pups_df)

pups_df

```

Use absolute path. 

```{r, eval = FALSE}

pups_df = read_csv("~/Documents/School/Fall2024/BIST P8105/data_wrangling_I/data/FAS_pups.csv")


```

## Look at read_csv options 

col_names and skipping rows 

```{r}
litters_df = 
  read_csv(
    file="data/FAS_litters.csv",
    col_names = FALSE, 
  )
spec(litters_df)
show_col_types = FALSE
```

What about missing data 
```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", ".")
  )

litters_df = janitor::clean_names(litters_df)

pull(litters_df, gd0_weight)
```

What if we code `group` as a factor variable?

```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", "."), 
    col_types = cols(
      Group = col_factor()
    )
  )
```


## Importing an excel file 

Import MLB 2011 summary data

```{r}
mlb_df = read_excel("data/mlb11.xlsx", sheet = "mlb11")
```


Import SAS data 
```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
```

## Never use read.csv()

```{r}
litter_df = read_csv("data/FAS_litters.csv")
```


## Never do this either: 
```{r}
litters_df$L

```


